Ripple Protocol Consensus Algorithm Review

### 2015-05-11 paper.tex
close to finished

    \item Reviewed whitepapers and development documentation at \url{https://ripple.com}.

@@

Ripple has diverged significantly from the original
concept\cite{btcmag-introducing-ripple} of a decentralized network recording
the transferral of money explicitly as debt relationships between parties.
(Figure \ref{fig:orig-ripple}, showing entities A, B, and C participating in a
transaction) While the concept of a debt relationship still exists in the form
of trustlines between participants, the bulk of the codebase and developer
documentation now focuses\footnote{For instance the ``Tutorials'' section of
the Ripple website only explains how to create transactions that modify the
ledger; there is almost no information available on how to actually use the
trustlines feature.} on the use and maintenance of a global ledger of
transactions and account balances.  Additionally a native currency, XRP, has
been added, which is used for anti-spam transaction fees\footnote{A small
amount of XRP is irrovocably destroyed for every transaction.} and to
serve as a universal currency. (Figure \ref{fig:ripple-labs-ripple})

@@

Authentication is performed with a simple signature scheme; a scripting system
is not available, nor is multisignature support. To implement the latter a
highly complex single-purpose scheme\cite{ripple-wiki-multisign} has been
proposed.

@@

Increasing the connectivity past the $20\%$ threshold, as shown in Figure
\ref{fig:unl-clique-consensus}, brings both cliques into agreement. However
what happens to the blockchain? The Ripple Protocol Consensus Algorithm paper
is unclear on this point. For instance, the loss and restoration of
connectivity may have been associated with network latency. The paper does
state that "nodes who’s latency grows larger than a preset bound $b$ are
removed from all UNLs"\cite[3.4.1]{ripple-consensus-paper} but does not say if
this removal is meant to be permanent\footnote{In the existing implementation
restarting the node does reload the UNL set from the config file.} nor if by
``all'' they refer to removal from UNLs on all nodes in the network.
Additionally while the author did not have sufficient time to investigate this
point fully, the source code doesn't appear to have any provision for
reorganizations of the ledger when a longer fork is detected.

@@

    \begin{tabular}{cc|c|c|c}
        \multicolumn{2}{c}{} & \multicolumn{3}{c}{Majority} \\
        \multicolumn{2}{c}{} & $100\%$ & $50\%$ & $10\%$ \\ \cline{3-5}
        \multirow{3}{*}{Local} & $100\%$ & Consensus & DoS & DoS \\ \cline{2-5}
        ~ & $50\%$ & DoS & DoS & Doublespend \\ \cline{2-5}
        ~ & $10\%$ & Doublespend & Doublespend & Doublespend \\
    \end{tabular}

@@

\section{Attack scenarios}

@@

The attacker creates large numbers of transactions - either covertely or
overtly - until the Ripple network fails due to overload, or more likely,
Ripple users are priced out of their ability to use the Ripple network.
Transactions pay a minimum fee per transaction by destroying XRP; the fee per
transaction is set by a vote between all validators on the UNL list. In the
short to medium term such an attack would quickly drive up that fee. A
succesful attack requires deep pockets as the attacker must outspend other
Ripple users.

@@

    \item A clever attacker wishing to overtly attack the Ripple network can
        masquerade their traffic as legit economic activity; from the point of
        view of Ripple Labs - a major XRP owner - such an attack may not even
        be considered an attack!\footnote{Similar to the debates in the Bitcoin
        ecosystem about whether or not fee-paying uses of the Bitcoin
        blockchain for non-Bitcoin-denominated transactions constitute
        ``attacks''.}

    \item Attacking the poor scalability of Ripple is highly synergistic with
        attacks depending on the high centralization of the Ripple network as
        even ``unsuccesful'' scalability attacks drive further centralization.
        For instance, if the response to a scalability attack is for Ripple
        Labs to beef up the validation nodes they control, and users stop
        running full nodes themselves, the network is more vulnerable to any
        attacks on those centrally managed validation nodes.

@@

\subsection{Coercion of validators}

@@

Validators are coerced through poltical/legal/criminal means into attacking
some or all Ripple users by preventing valid transactions from entering the
ledger (DoS) and/or by allowing invalid transactions to enter the ledger.
(theft) The cost estimates above are very rough, and assuming the attack is via
legal means; the cost may potentially be as low as zero in some scenarios like
blackmail. In preparation for this attack the attackers may also attempt to
reduce the number of validators by with, for instance, a transaction flood
attack.

@@

    \item The Ripple protocol is constructed such that it is possible to create
        many, although not all, types of fraud
        proofs\cite{irc-btcdev-gmaxwell-fraud-proof, btctalk-pkt-fraud-proofs}
        required to prove an invalid ledger. While the software does not
        currently implement fraud proofs, this is functionality that could be
        added in the future without a change to the consensus. If widely used
        fraud proofs could prevent some types of coercion attacks that force
        validators to create invalid ledgers by making it likely for the result
        of the attack to be DoS rather than profitable theft. In short, the
        attack would simply shut the Ripple network down for all users rather
        than result in a succesful theft.

    \item Some people would argue that the recently announced
        plans\cite{cointelegraph-ripple-aml} to require verified user
        identification to use the Ripple network in the near future are an
        example of a succesful coercion attack. It can be argued that Ripple
        Labs is increasing the risk of coercion attacks by simultaneously
        adopting a ``regulator friendly'' position and promoting the use of
        Ripple across many different jurisdictions.

@@

\subsection{Software backdoor}

@@

    \item \textbf{Currently Ripple Labs does not provide a secure way to
        download any of their software.} While binaries are available, the
        mechanism to download them is a Ubuntu package repository over insecure
        HTTP; while source code is available through git, neither commits nor
        release tags are signed in anyway. \textbf{This is a serious ommission
        that has lead to significant monetary losses in the past.} Ripple Labs
        should be following industry best-practice by signing git commits and
        tags\cite{wladimir-git-pgp} as well as PGP signing their Ubuntu
        packages.

@@

\subsection{Theft of validator secret keys}

@@

The attacker steals illegitimate control of the secret keys used by validators,
particularly the validators in the default UNL. The theft may happen via an
exploit, backdoor, compromised hosting company, etc. While not useful for the
attacker in of itself, the stolen keys can be used for further atttacks such as
DoS or a simulated ledger attack.

@@

    \item Due to the high centralization of UNL validators, a relatively small
        number of validators need to be compromised for a succesful attack.

    \item Most validators are running nearly the same, if not the exact same,
        software, making it likely for backdoor/exploit attack to succesfully
        exfiltrate the secret keys. Due to the high consensus-critical
        complexity of the Ripple protocol and lack of a consensus library
        alternate validator implementations are currently infeasible, making it
        difficult to change this situation.

@@

\subsection{Simulated ledger}

@@

An attacker with the ability to create signatures from nodes on the UNL can
create an entirely simulated ledger in parallel to the ``official'' Ripple
ledger, containing any transaction the attacker wishes. In conjunction with
sybil attacking the target the attacker can then perform double-spend attacks.
Sybil attacks are relatively cheap, requiring just the aquisition of a large
number of IP addresses, e.g. via cloud hosting and/or botnet rental. In some
situations sybil attacks can even have zero cost, such as the internal IT staff
within the target with access to network infrastructure, the target's ISP, or
government actors with MITM capabilities such as the
NSA.\cite{schneier-nsa-mitm}

@@

\section{Attack flowchart}

@@

Let's go through a full scenario, involving multiple attacks at once.  Ripple
Labs has diversified their default UNL such that the $80\%$ of the validators
are run by publicly known institutions evenly spread across the following five
countries, United States, France, Switzerland, Russia, and China, with the
remaining fifth being controlled by pseudoanonymous ``DarkWeb'' entities like
The Silk Road Five, and Agora Market. As any one entity only comprises at most
$1/6\text{th}$ of the UNL no single actor can block transactions.

@@

At this point Russia with $20\%$ UNL control could simply vote to freeze
Shell's assets, achieving a denial-of-service. However, it's decided it would
be better if the funds were actually recovered instead.

Russia sets up a fake Chinese oil importer, hiring another Chinese company to
do the actual transportation, and has them contact Shell asking to purchase
some oil. They also find a wealthy Chinese investor with \$100 million USD
worth of yuan in an account with a major Chinese bank, and has that investor
``invest'' in the new importer. Meanwhile Russia's cyberwarfare team gets to
work, succesfully compromising the Swiss and French validation nodes, and
using the fact that they have agents in major Ukrainian ISPs to sybil attack
the Ukrainian-based Shell office actually handling the sale. They can't
compromise the US nodes - they're on a newer version - but they do happen to
find a new zero-day consensus bug in the newer version.

@@

\section{Conclusions and future work}

@@

A key question that should be answered in future work is if the goals of the
Ripple system need global consensus at all? If global consensus can be avoided,
or at least it's use minimized, many of these issues may go away.

>>>>

    \item Reviewed white papers and development documentation at \url{https://ripple.com}.

@@

Ripple has diverged significantly from the original
concept\cite{btcmag-introducing-ripple} of a decentralized network recording
the transferral of money explicitly as debt relationships between parties.
(Figure \ref{fig:orig-ripple}, showing entities A, B, and C participating in a
transaction) While the concept of a debt relationship still exists in the form
of trustlines between participants, the bulk of the codebase and developer
documentation now focuses\footnote{For instance the ``Tutorials'' section of
the Ripple website only explains how to create transactions that modify the
ledger; there is almost no information available on how to actually use the
trustlines feature.} on the use and maintenance of a global ledger of
transactions and account balances.  Additionally a native currency, XRP, has
been added, which is used for anti-spam transaction fees\footnote{A small
amount of XRP is irrevocably destroyed for every transaction.} and to
serve as a universal currency. (Figure \ref{fig:ripple-labs-ripple})

@@

Authentication is performed with a simple signature scheme; a scripting system
is not available, nor is multi-signature support. To implement the latter a
highly complex single-purpose scheme\cite{ripple-wiki-multisign} has been
proposed.

@@

Increasing the connectivity past the $20\%$ threshold, as shown in Figure
\ref{fig:unl-clique-consensus}, brings both cliques into agreement. However
what happens to the blockchain? The Ripple Protocol Consensus Algorithm paper
is unclear on this point. For instance, the loss and restoration of
connectivity may have been associated with network latency. The paper does
state that ``nodes whose latency grows larger than a preset bound $b$ are
removed from all UNLs''\cite[3.4.1]{ripple-consensus-paper} but does not say if
this removal is meant to be permanent\footnote{In the existing implementation
restarting the node does reload the UNL set from the config file.} nor if by
``all'' they refer to removal from UNLs on all nodes in the network.
Additionally while the author did not have sufficient time to investigate this
point fully, the source code doesn't appear to have any provision for
reorganizations of the ledger when a longer fork is detected.

@@

    \begin{tabular}{cc|c|c|c}
        \multicolumn{2}{c}{} & \multicolumn{3}{c}{Majority} \\
        \multicolumn{2}{c}{} & $100\%$ & $50\%$ & $10\%$ \\ \cline{3-5}
        \multirow{3}{*}{Local} & $100\%$ & Consensus & DoS & DoS \\ \cline{2-5}
        ~ & $50\%$ & DoS & DoS & Double-spend \\ \cline{2-5}
        ~ & $10\%$ & Double-spend & Double-spend & Double-spend \\
    \end{tabular}

@@

\section{Attack Scenarios}

@@

The attacker creates large numbers of transactions - either covertly or
overtly - until the Ripple network fails due to overload, or more likely,
Ripple users are priced out of their ability to use the Ripple network.
Transactions pay a minimum fee per transaction by destroying XRP; the fee per
transaction is set by a vote between all validators on the UNL list. In the
short to medium term such an attack would quickly drive up that fee. A
successful attack requires deep pockets as the attacker must out-spend other
Ripple users.

@@

    \item A clever attacker wishing to overtly attack the Ripple network can
        masquerade their traffic as legit economic activity. From the point of
        view of Ripple Labs - a major XRP owner - such an attack may not even
        be considered an attack!\footnote{Similar to the debates in the Bitcoin
        ecosystem about whether or not fee-paying uses of the Bitcoin
        blockchain for non-Bitcoin-denominated transactions constitute
        ``attacks''.}

    \item Attacking the poor scalability of Ripple is highly synergistic with
        attacks depending on the high centralization of the Ripple network as
        even ``unsuccessful'' scalability attacks drive further centralization.
        For instance, if the response to a scalability attack is for Ripple
        Labs to beef up the validation nodes they control, and users stop
        running full nodes themselves, the network is more vulnerable to any
        attacks on those centrally managed validation nodes.

@@

\subsection{Coercion of Validators}

@@

Validators are coerced through political/legal/criminal means into attacking
some or all Ripple users by preventing valid transactions from entering the
ledger (DoS) and/or by allowing invalid transactions to enter the ledger.
(theft) The cost estimates above are very rough, and assuming the attack is via
legal means; the cost may potentially be as low as zero in some scenarios like
blackmail. In preparation for this attack the attackers may also attempt to
reduce the number of validators by with, for instance, a transaction flood
attack.

@@

    \item The Ripple protocol is constructed such that it is possible to create
        many, although not all, types of fraud
        proofs\cite{irc-btcdev-gmaxwell-fraud-proof, btctalk-pkt-fraud-proofs}
        required to prove an invalid ledger. While the software does not
        currently implement fraud proofs, this is functionality that could be
        added in the future without a change to the consensus. If widely used
        fraud proofs could prevent some types of coercion attacks that force
        validators to create invalid ledgers by making it likely for the result
        of the attack to be DoS rather than profitable theft. In short, the
        attack would simply shut the Ripple network down for all users rather
        than result in a successful theft.

    \item Some people would argue that the recently announced
        plans\cite{cointelegraph-ripple-aml} to require verified user
        identification to use the Ripple network in the near future are an
        example of a successful coercion attack. It can be argued that Ripple
        Labs is increasing the risk of coercion attacks by simultaneously
        adopting a ``regulator friendly'' position and promoting the use of
        Ripple across many different jurisdictions.

@@

\subsection{Software Backdoor}

@@

    \item \textbf{Currently Ripple Labs does not provide a secure way to
        download any of their software.} While binaries are available, the
        mechanism to download them is a Ubuntu package repository over insecure
        HTTP; while source code is available through git, neither commits nor
        release tags are signed in anyway. \textbf{This is a serious omission
        that has lead to significant monetary losses in the past.} Ripple Labs
        should be following industry best-practice by signing git commits and
        tags\cite{wladimir-git-pgp} as well as PGP signing their Ubuntu
        packages.

@@

\subsection{Theft of Validator Secret Keys}

@@

The attacker steals illegitimate control of the secret keys used by validators,
particularly the validators in the default UNL. The theft may happen via an
exploit, backdoor, compromised hosting company, etc. While not useful for the
attacker in of itself, the stolen keys can be used for further attacks such as
DoS or a simulated ledger attack.

@@

    \item Due to the high centralization of UNL validators, a relatively small
        number of validators need to be compromised for a successful attack.

    \item Most validators are running nearly the same, if not the exact same,
        software, making it likely for backdoor/exploit attack to successfully
        exfiltrate the secret keys. Due to the high consensus-critical
        complexity of the Ripple protocol and lack of a consensus library
        alternate validator implementations are currently infeasible, making it
        difficult to change this situation.

@@

\subsection{Simulated Ledger}

@@

An attacker with the ability to create signatures from nodes on the UNL can
create an entirely simulated ledger in parallel to the ``official'' Ripple
ledger, containing any transaction the attacker wishes. In conjunction with
sybil attacking the target the attacker can then perform double-spend attacks.
Sybil attacks are relatively cheap, requiring just the acquisition of a large
number of IP addresses, e.g. via cloud hosting and/or botnet rental. In some
situations sybil attacks can even have zero cost, such as the internal IT staff
within the target with access to network infrastructure, the target's ISP, or
government actors with MITM capabilities such as the
NSA.\cite{schneier-nsa-mitm}

@@

\section{Attack Flowchart}

@@

Let's go through a full scenario involving multiple attacks at once.  Ripple
Labs has diversified their default UNL such that the $80\%$ of the validators
are run by publicly known institutions evenly spread across the following five
countries, United States, France, Switzerland, Russia, and China, with the
remaining fifth being controlled by pseudoanonymous ``DarkWeb'' entities like
The Silk Road Five, and Agora Market. As any one entity only comprises at most
$1/6\text{th}$ of the UNL no single actor can block transactions.

@@

At this point Russia's $20\%$ UNL control could simply vote to freeze Shell's
assets, achieving a denial-of-service. However, it's decided it would be better
if the funds were actually recovered instead.

Russia sets up a fake Chinese oil importer, hiring another Chinese company to
do the actual transportation, and has them contact Shell asking to purchase
some oil. They also find a wealthy Chinese investor with \$100 million USD
worth of yuan in an account with a major Chinese bank, and has that investor
``invest'' in the new importer. Meanwhile Russia's cyberwarfare team gets to
work, successfully compromising the Swiss and French validation nodes, and
using the fact that they have agents in major Ukrainian ISPs to sybil attack
the Ukrainian-based Shell office actually handling the sale. They can't
compromise the US nodes - they're on a newer version - but they do happen to
find a new zero-day consensus bug in the newer version.

@@

\section{Conclusions and Future Work}

@@

A key question that should be answered in future work is if the goals of the
Ripple system need global consensus at all? If global consensus can be avoided,
or at least its use minimized, many of these issues may go away.
